---
title: "RIOULT 2016 Audience Survey Findings"
author: "Tim Boomer"
date: "August 1, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = FALSE, results = 'hide', 
                      warning = FALSE, message = FALSE)
```

## Overview
An audience survey was conducted at each of the non-gala season performances during
June, 2016. The survey design built off the questions on a 2013 survey to
facilitate period to period comparisons of audience demographics. Results from 394 2016 surveys were tabulated representing a 20% response rate. This report summarizes key findings.

```{r prep}
library(readxl)
library(readr)
library(stringr)
# library(plyr)
library(dplyr)
library(reshape2)
library(ggmap)

setwd("C:/Users/tboom_000/Documents/Personal/Projects/RioultSurvey")
source('./script/remap_char.R') # Function to clean up dance co names.

# Function to map zip codes to state
map_state <- function(x) {
     prefix <- as.numeric(substr(x, 1, 3))
     if(is.na(prefix))                      {state <- NA}
     else if(prefix >= 10 & prefix <= 27)   {state  <- "massachusetts"}
     else if(prefix >= 60 & prefix <= 69)   {state  <- "connecticut"}
     else if(prefix >= 70 & prefix <= 89)   {state  <- "new jersey"}
     else if(prefix >= 100 & prefix <= 149) {state  <- "new york"}
     else                                   {state  <- "Other"}
     return(state)
}

as_binary <- function(x) {
     x[x == 'x'] <- 1 # For 2013 encoded data
     x[x == 'T'] <- 1 # For 2016 encoded data
     x <- as.integer(x)
     x[is.na(x)] <- 0
     return(x)
}

# Load and clean data
# Load data from Excel
types <- c(rep('text', 26), 'numeric', rep('text', 29))
data_2016 <- read_excel('./SourceData/SurveyResults2016import.xls', 
                        col_types = types, sheet = 1, skip = 9)

data_2016 <- data_2016[1:394, c(1, 5:55)] # Filter out unneeded cols & blank rows

colnames(data_2016) <- c('Performance', 'Rioult_Direct', 'Joyce_Direct', 
                         'Any_Print_Ad', 'Times', 'TimeOut', 'MetroNY', 
                         'AMNY_Ad', 'Article_Blog', 'Any_Social_Media', 'Facebook',
                         'Twitter', 'Instagram', 'Personal_Recommendation', 
                         'Other_Source', 'Source_Notes', 'Number_Performances',
                         'Like_Company', 'Like_Specific_Dances', 'Like_Modern_Dance', 
                         'Like_Dance', 'Like_Music', 'Like_Joyce', 'Joyce_Years_Member', 
                         'Am_Dancer', 'Am_FriendOrFamily', 'Am_Guest', 'Other_Reason',
                         '92Y', 'BAM', 'Dixon', 'Pillow',
                         'Judson', 'Joyce', 'LincolnCtr',
                         'CityCtr', 'NYLiveArts', 'Triskelion',
                         'Other_Venue', 'Dance_Company1', 'Dance_Company2',
                         'Dance_Company3', 'Interested_InSchool',
                         'Interested_Children', 'Interested_Pre-Prof',
                         'Interested_Adult', 'Gender', 'Zip_Code', 'Age', 'Ethnicity',
                         'Income', 'Education')

# Modify variable types
data_2016[, c(2:15, 18:23, 25:27, 43:46)] <- 
     lapply(data_2016[, c(2:15, 18:23, 25:27, 43:46)], as_binary)
data_2016$Joyce_Years_Member <- as.numeric(data_2016$Joyce_Years_Member)
data_2016[, c(1, 17, 29:38, 47, 49:52)] <- 
     lapply(data_2016[, c(1, 17, 29:38, 47, 49:52)], as.factor)
data_2016[, 28] <- as.integer(!is.na(data_2016[,28])) # Field contains text and "T"

# Resequence factor levels
data_2016$Income <- factor(data_2016$Income, 
                           levels = c("<25", "25-50", "50-75", "75-100",
                                      "100-150", "150-200", "200+"))
data_2016$Number_Performances <- factor(data_2016$Number_Performances,
                                        levels = c("First time", "1-4 times", "5+ times"))
data_2016$Education <- factor(data_2016$Education, 
                              levels = c("Some high school", "Some college",
                                         "College grad", "Some graduate school",
                                         "Grad/prof degree"))
data_2016$Ethnicity <- relevel(data_2016$Ethnicity, "White")

# Change "Any_" variables so they are true if any subcategory is true
data_2016$Any_Print_Ad = apply(data_2016[, 4:9], 1, sum)
data_2016$Any_Social_Media = apply(data_2016[, 10:13], 1, sum)

# Create Program variable
data_2016$Program = as.factor(substr(data_2016$Performance, 9,9))

# Create text file of unique dance company names for mapping
# !!! Only run when there is a change in the table !!! Only run when there is a change in the table
# company <- sort(unique(c(data_2016$Dance_Company1, data_2016$Dance_Company2, data_2016$Dance_Company3)))
# write(company, file = './cache/dance_co.txt')

# Remap Dance Company names
map_table <- read_csv('./cache/dance_co_mapping.csv')
data_2016$Dance_Company1_norm <- map_char(data_2016$Dance_Company1, map_table, "Input", "Mapped_To")
data_2016$Dance_Company2_norm <- map_char(data_2016$Dance_Company2, map_table, "Input", "Mapped_To")
data_2016$Dance_Company3_norm <- map_char(data_2016$Dance_Company3, map_table, "Input", "Mapped_To")

# Aggregate data by zip code and map zip codes to state
data_2016$State <- sapply(data_2016$Zip_Code, map_state)
zip_count_2016 <- data_2016 %>% group_by(region = Zip_Code) %>% summarize(value = n())
state_count_2016 <- data_2016 %>% group_by(region = State) %>% summarize(value = n())


#-------------------------------------------------------------------------------
# 2013 Data
#-------------------------------------------------------------------------------
data_2013 <- read_excel('./SourceData/SurveyImport2013.xls', sheet = 2)
# Normalize zip codes
data_2013$Zip <- str_replace(data_2013$Zip, ".000000", "")
# Filter out columns not required for period to period comparison
data_2013 <- data_2013[, c(12,33:38)] 
colnames(data_2013) <- c('Number_Performances','Gender', 'Zip_Code', 'Age', 
                         'Ethnicity', 'Income', 'Education')

# Modify variable types
data_2013 <- data.frame(sapply(data_2013, function(x) gsub("^$|^ $", NA, x)))
data_2013 <- data.frame(sapply(data_2013, as.factor))

# Rename and resequence factor levels
levels(data_2013$Age) <- c("<18", "18-24", "25-29", "30-44", "45-64", "45-64", "65+")

levels(data_2013$Income) <- c("<25", "100-150", "150-200", "200+", "25-50", "50-75", "75-100")
data_2013$Income <- factor(data_2013$Income, 
                           levels = c("<25", "25-50", "50-75", "75-100",
                                      "100-150", "150-200", "200+"))

levels(data_2013$Number_Performances) <- c("1-4 times", "1-4 times", "5+ times", "First time")
data_2013$Number_Performances <- factor(data_2013$Number_Performances,
                                        levels = c("First time", "1-4 times", "5+ times"))

levels(data_2013$Education) <- c("College grad", "Grad/prof degree", "Grad/prof degree",
                                 "Grad/prof degree", "Some high school", "Some college",
                                 "Some college", "Some graduate school", "Some graduate school",
                                 "Some high school", "Some high school")
data_2013$Education <- factor(data_2013$Education, 
                              levels = c("Some high school", "Some college",
                                         "College grad", "Some graduate school",
                                         "Grad/prof degree"))

data_2013$Ethnicity[data_2013$Ethnicity == "Prefer not to say"] <- NA
levels(data_2013$Ethnicity) <- c("Afro-American", "Am Indian", "Asian", "White",
                                 "Hispanic", "Multiracial", "Multiracial",
                                 "White", "White")
data_2013$Ethnicity <- relevel(data_2013$Ethnicity, "White")


# Combine data with 2016
data_combined <- cbind(Year = "2013", data_2013)
temp_2016 <- cbind(Year = "2016", data_2016)
temp_2016 <- select(temp_2016, c(18, 48:53))
temp_2016 <- cbind(Year = "2016", temp_2016)
data_combined <- rbind(data_combined, temp_2016)
rm(temp_2016)

```


## Audience Demographics
Of the total 391 of 2016 survey respondents to the question, "Have you ever seen the company before?" 
over half (55%) indicated they were attending a RIOULT performance for the first time. This was down from 65% of the 455 responses to the question in 2013.

``` {r fig.width = 6, fig.height = 3, fig.align = 'center'}
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(RColorBrewer)

# Standard Plot Parameters
pal_class <- scale_fill_brewer(type = "qual", palette = 'Pastel1')
pal_seq <- scale_fill_brewer(type = "seq", palette = 'BuGn')
n_text_size = 3

# Period to Period Comparisons
# Number of Performances
plot_data <-  group_by(data_combined, Year, Number_Performances) %>%
     filter(!is.na(Number_Performances)) %>%
     summarize(n = n(), na.exclude = TRUE) %>%
     mutate(Percent = 100* n / sum(n))
response_count <- group_by(plot_data, Year) %>% summarize(n = sum(n))
ggplot(plot_data, aes(x = Year, y = Percent, group = Number_Performances, fill = Number_Performances)) + 
     geom_bar(stat = "identity", width = 0.5) +
     pal_seq +
     ggtitle("Number of RIOULT Performances by Survey Period") +
     annotate("text", label = paste("N =", as.character(response_count$n)), x = 1:2, y = 105, 
              size = 2)
```

\pagebreak

A period to period comparison of audience demographics shows the RIOULT audience is on average affluent and well-educated. Twenty-two percent of respondents have self-reported income over $200,000+ and over half have a graduate or professional degree. Compared to the 2013 survey the audience is grayer with almost 40% of 2016 respondents indicating their age was 65+.  In addition, 2016 respondents tended to to have higher income and a higher level of education.

``` {r fig.width = 3, fig.height = 3}
# Age
plot_data <-  group_by(data_combined, Year, Age) %>%
     filter(!is.na(Age)) %>%
     summarize(n = n(), na.exclude = TRUE) %>%
     mutate(Percent = 100* n / sum(n))
response_count <- group_by(plot_data, Year) %>% summarize(n = sum(n))
ggplot(plot_data, aes(x = Year, y = Percent, group = Age, fill = Age)) + 
     geom_bar(stat = "identity", width = 0.5) +
     pal_seq +
     ggtitle("Age Distribution") +
     annotate("text", label = paste("N =", as.character(response_count$n)), x = 1:2, y = 105, 
              size = 2)

# Income
plot_data <-  group_by(data_combined, Year, Income) %>%
     filter(!is.na(Income)) %>%
     summarize(n = n(), na.exclude = TRUE) %>%
     mutate(Percent = 100* n / sum(n))
response_count <- group_by(plot_data, Year) %>% summarize(n = sum(n))
ggplot(plot_data, aes(x = Year, y = Percent, group = Income, fill = Income)) + 
     geom_bar(stat = "identity", width = 0.5) +
     pal_seq +
     ggtitle("Income") + 
     annotate("text", label = paste("N =", as.character(response_count$n)), x = 1:2, y = 105, 
              size = 2)
```
```{r fig.width = 4, fig.height = 3, fig.align = 'center'}
# Education
plot_data <-  
     group_by(data_combined, Year, Education) %>%
     filter(!is.na(Education)) %>%
     summarize(n = n(), na.exclude = TRUE) %>%
     mutate(Percent = 100* n / sum(n))
response_count <- group_by(plot_data, Year) %>% summarize(n = sum(n))
ggplot(plot_data, aes(x = Year, y = Percent, group = Education, fill = Education)) + 
     geom_bar(stat = "identity", width = 0.5) +
     pal_seq +
     ggtitle("Education") +
     annotate("text", label = paste("N =", as.character(response_count$n)), x = 1:2, y = 105, 
              size = 2)
```

\pagebreak

## Demographics by 2016 Season Program

Two programs were presented during the season. Program A included
live music, featured Kathleen Turner as narrator, and consisted of three works
thematically connected by the Trojan war. Program B was a combination of
new and old works. Surveys were collected at four Program A performances (232 survey forms or 21% response rate) and three Program B performances (162 survey forms or 19% response rate).

Respondents who attended program B were on the whole older and had a higher level of education than those who attended program A. However, program A attendees tended to have higher overall income. 

``` {r fig.width = 3, fig.height = 3}
plot_data <-  group_by(data_2016, Program, Age) %>%
     filter(!is.na(Age)) %>%
     summarize(n = n(), na.exclude = TRUE) %>%
     mutate(Percent = 100* n / sum(n))
response_count <- group_by(plot_data, Program) %>% summarize(n = sum(n))
ggplot(plot_data, aes(x = Program, y = Percent, fill = Age)) + 
     geom_bar(stat = "identity", width = 0.5) + 
     pal_seq +
     ggtitle("Age by Program") +
     annotate("text", label = paste("N =", as.character(response_count$n)), x = 1:2, y = 105, 
              size = 2)

plot_data <-  group_by(data_2016, Program, Income) %>%
     filter(!is.na(Income)) %>%
     summarize(n = n(), na.exclude = TRUE) %>%
     mutate(Percent = 100* n / sum(n))
response_count <- group_by(plot_data, Program) %>% summarize(n = sum(n))
ggplot(plot_data, aes(x = Program, y = Percent, fill = Income)) + 
     geom_bar(stat = "identity", width = 0.5) +
     pal_seq +
     ggtitle("Income by Program") +
     annotate("text", label = paste("N =", as.character(response_count$n)), x = 1:2, y = 105, 
              size = 2)
```
```{r fig.width = 4, fig.height = 3, fig.align = 'center'}
plot_data <-  group_by(data_2016, Program, Education) %>%
     filter(!is.na(Education)) %>%
     summarize(n = n(), na.exclude = TRUE) %>%
     mutate(Percent = 100* n / sum(n))
response_count <- group_by(plot_data, Program) %>% summarize(n = sum(n))
ggplot(plot_data, aes(x = Program, y = Percent, fill = Education)) + 
     geom_bar(stat = "identity", width = 0.5) +
     pal_seq +
     ggtitle("Education by Program") +
     annotate("text", label = paste("N =", as.character(response_count$n)), x = 1:2, y = 105, 
              size = 2)
```

\pagebreak
Over 2/3 of respondents at both programs were Women and over 3/4 were white.
Program A respondents had a greater proportion of men but showed less ethnic diversity than 
program B.

``` {r fig.width = 3, fig.height = 3}
plot_data <-  group_by(data_2016, Program, Gender) %>%
     filter(!is.na(Gender)) %>%
     summarize(n = n(), na.exclude = TRUE) %>%
     mutate(percent = 100 * n / sum(n))
response_count <- group_by(plot_data, Program) %>% summarize(n = sum(n))
ggplot(plot_data, aes(x = Program, y = percent, fill = Gender)) + 
     geom_bar(stat = "identity", width = 0.5) +
     pal_class +
     ggtitle("Gender by Program") +
     annotate("text", label = paste("N =", as.character(response_count$n)), x = 1:2, y = 105, 
              size = 2)

plot_data <-  group_by(data_2016, Program, Ethnicity) %>%
     filter(!is.na(Ethnicity)) %>%
     summarize(n = n(), na.exclude = TRUE) %>%
     mutate(Percent = 100* n / sum(n))
response_count <- group_by(plot_data, Program) %>% summarize(n = sum(n))
ggplot(plot_data, aes(x = Program, y = Percent, fill = Ethnicity)) + 
     geom_bar(stat = "identity", width = 0.5) +
     pal_class +
     ggtitle("Ethnicity by Program") +
     annotate("text", label = paste("N =", as.character(response_count$n)), x = 1:2, y = 105, 
              size = 2)
```

However, both programs attracted approximately the same proportion of new audience 
(those who are attending their first performance) versus respondents who had 
previously attended RIOULT performances.

``` {r fig.width = 6, fig.height = 3, fig.align = 'center'}
# Number of Performances by Program
plot_data <-  group_by(data_2016, Program, Number_Performances) %>%
     filter(!is.na(Number_Performances)) %>%
     summarize(n = n(), na.exclude = TRUE) %>%
     mutate(Percent = 100* n / sum(n))
response_count <- group_by(plot_data, Program) %>% summarize(n = sum(n))
ggplot(plot_data, aes(x = Program, y = Percent, fill = Number_Performances)) + 
     geom_bar(stat = "identity", width = 0.5) +
     pal_seq +
     ggtitle("Number of RIOULT Performances by Program") +
     annotate("text", label = paste("N =", as.character(response_count$n)), x = 1:2, y = 105, 
              size = 2)
```

\pagebreak

## Marketing and Promotion

In response to "How did you hear about the performance?" Personal Recommendation
and the Joyce Theater Flyer/Website were the two most frequent mentions. Of respondents
who selected "Other", 41% wrote-in TDF.

``` {r fig.width = 6, fig.height = 3, fig.align = 'center'}
plot_data <- melt(data_2016[, 2:15]) %>% group_by(Source = variable) %>% 
     summarize(Mentions = sum(value))
# Plot Sources
# # Source of Awareness
plot1 <- filter(plot_data, as.integer(Source) %in% c(1:3, 8, 9, 13:14))
plot1$Source <- factor(plot1$Source, 
                       levels = plot1$Source[order(plot1$Mentions)])
ggplot(plot1, aes(x = Source, y = Mentions)) + 
     geom_bar(stat = "identity") +
     coord_flip() +
     ggtitle("Source of Awareness: Number of Mentions")

```



Respondents who selected Advertising or Social Media as a source had the option of providing a specific source. Out of the total responses that cited advertising as a source, the New York Times and TimeOut ads were most frequently mentioned. Facebook was the most frequently mentioned social medium.

``` {r fig.width = 3, fig.height = 3}
# Plot Ad Detail
plot2 <- filter(plot_data, as.integer(Source) %in% 4:7)
ggplot(plot2, aes(x = Source, y = Mentions)) + 
     geom_bar(stat = "identity", width = 0.5) +
     # coord_flip() +
     ggtitle("Advertising Detail")

# Plot Social Media Detail
plot3 <- filter(plot_data, as.integer(Source) %in% 10:12)
ggplot(plot3, aes(x = Source, y = Mentions)) + 
     geom_bar(stat = "identity", width = 0.5) +
     ggtitle("Social Media Detail")
```

\pagebreak

In response to "What helped you decide to attend this performance?" respondents could
select from a set of "I like" and a set of "I am" choices. Of the "I like" choices, Dance, Modern Dance, the Joyce, and the Company (RIOULT) were most frequently cited. "Dancer" and "Friend or Family" were the most frequently cited "I am" selections.

```{r fig.width = 6, fig.height = 3, fig.align = 'center'}
plot_data <- melt(data_2016[, c(18:23, 25:28)]) %>% 
     group_by(Reason = variable) %>% 
     summarize(Mentions = sum(value))
plot_data$Reason <- factor(plot_data$Reason, 
                           levels = plot_data$Reason[order(plot_data$Mentions)])
ggplot(plot_data, aes(x = Reason, y = Mentions)) + 
     geom_bar(stat = "identity") +
     pal_seq +
     coord_flip() +
     ggtitle("Reason for Attending")


joyce_members <- sum(!is.na(data_2016$Joyce_Years_Member))
avg_years <- median(data_2016$Joyce_Years_Member, na.rm = TRUE)
```

Joyce members were asked to indicate the number of years they had been members. Of the `r joyce_members` respondents, the average number of years was `r round(avg_years, 1)`.

## Dance Companies and Venues
Respondents were asked to write in their top three favorite dance companies other
than RIOULT. Respondents wrote in over 135 companies, of which Alvin Ailey, 
NYC Ballet, and ABT were most frequently mentioned. The full list is in the appendix.

``` {r fig.width = 6, fig.height = 3, fig.align = 'center'}
plot_data <- sort(table(unlist(data_2016[,54:56])), decreasing = FALSE)
plot_data <- filter(as.data.frame(plot_data), Freq > 7)
names(plot_data) <- c("Company", "Mentions")
ggplot(plot_data, aes(x = Company, y = Mentions)) + 
     geom_bar(stat = "identity") +
     coord_flip() +
     ggtitle("Dance Company Mentions: Total")
```

\pagebreak

Respondents were asked how frequently they had attended performances at selected 
dance venues in the previous 12 months. The Joyce and Lincoln Center had the most mentions
and many respondents attended 5+ performances at these venues, suggesting a significant
portion of the RIOULT audience has subscriptions. City Center and BAM also had more than 50 mentions.

``` {r fig.width = 6, fig.height = 3, fig.align = 'center'}
plot_data <- gather(data_2016[,29:38]) %>%
     filter(!is.na(value)) %>%
     group_by(key, value) %>%
     summarize(n())
names(plot_data) <- c("Venue", "Performances", "Mentions")
venue_seq <- group_by(plot_data, Venue) %>%
     summarize(total = sum(Mentions)) %>%
     arrange(total)
plot_data$Venue <- factor(plot_data$Venue, levels = venue_seq$Venue, ordered = TRUE)
plot_data <- plot_data[order(plot_data$Venue, decreasing = FALSE), ]
ggplot(plot_data, aes(x = Venue, y = Mentions, fill = Performances)) +
     geom_bar(stat = "identity") +
     coord_flip() +
     pal_seq +
     ggtitle("Dance Venues Attended")
```

## Interest in Dance Education
Respondents were asked "Are you interested in RIOULT’s education/outreach programs?"
Forty-six expressed interest in adult programs and between 10 and 20 expressed interest
in programs for children and pre-professionals.

``` {r fig.width = 6, fig.height = 3, fig.align = 'center'}
plot_data <- melt(data_2016[, 43:46]) %>% group_by(Type = variable) %>%
summarize(Mentions = sum(value))
levels(plot_data$Type) <- c('In_School', 'Children', 'Pre-professional', 'Adult')
ggplot(plot_data, aes(x = Type, y = Mentions)) +
geom_bar(stat = "identity", width = 0.5) +
ggtitle("Interest in Dance Programs")
```

\pagebreak

## Geographic Composition
``` {r}
tri_state_pct <- sum(state_count_2016$value[state_count_2016$region %in% 
                                                 c('connecticut', 'new jersey', 'new york')]) /
     sum(state_count_2016$value[!is.na(state_count_2016$region)])
```

Of zip-codes provided by respondents, `r round(tri_state_pct * 100, 0)` percent were in the tri-state area.

``` {r fig.width = 5, fig.height = 3, fig.align = 'center'}
library(choroplethr)
library(choroplethrZip)
library(choroplethrMaps)
library(dplyr)

state_choropleth(state_count_2016,
                 title="Respondents by State: NY, NJ, CT",
                 legend="Number of Respondents",
                 zoom = c('connecticut', 'new jersey', 'new york'))
```

Within New York City, Manhattan residents were particularly concentrated on the Upper West Side and Chelsea. Brooklyn was also well represented, particularly Park Slope and Boerum Hill.

``` {r fig.width = 7, fig.height = 5, fig.align = 'center'}
nyc_fips = c(36005, 36047, 36061, 36081, 36085)

zip_choropleth(zip_count_2016, county_zoom = nyc_fips, num_colors = 7,
               msa_zoom="New York-Newark-Jersey City, NY-NJ-PA",
               title="Respondents by Zip Code",
               legend="Number of Respondents")
```

\pagebreak

## Appendix: Complete List of Other Dance Companies

``` {r results = 'asis'}
plot_data <- as.data.frame(sort(table(unlist(data_2016[,54:56])), decreasing = TRUE))
names(plot_data) <- c("Company", "Mentions")
knitr::kable(plot_data, caption = "Total Dance Company Mentions")
```

